---
title: "Lab 1"
author: "Branden Walter"
date: "`r Sys.Date()`"
output: html_document
---

# Question 1

a). Create a vector called countries which contains the names of 5 countries that you would like to visit some day.
```{r}
countries <- c("Germany", "Italy", "Russia", "Spain", "Romania")
```

b). Print out the elements of the vector.
```{r}
countries
```

c). Print out the 4th element of the vector.
```{r}
countries[4]
```

# Question 2
a). Create a vector called nums which contains the numbers 5, 6, 7, 8, 9,..., 199, 200 (consecutive integers)
```{r}
nums <- 5:200
```

b). Use the length() function to find the number of numbers in the vector.
```{r}
length(nums)
```

c). Use the sum() function to add up all of the values in nums.
```{r}
sum(nums)
```

d). Use the mean() function to find the average of all of the values in nums.
```{r}
mean(nums)
```

# Question 3
a). Create a vector called sleep which contains how many hours of sleep you get on each day of the week (starting with Monday).
```{r}
sleep <- c(8,8,8,8,8,8,8)
```

b). Create a vector called days which contains the names of the days of the week (starting with Monday).
```{r}
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```

c). Use the names() function to set the names of the sleep vector to the values of the days vector.
```{r}
names(sleep) <- days
```

d). Show two different ways to print out the number of hours of sleep you get on a Sunday.
```{r}
sleep
print(sleep)
```