---
title: 'Lab3: Data Frames'
author: "Branden Walter"
date: "`r Sys.Date()`"
output: html_document
---
# Question 0

#### (a) Delete all sample code from the R markdown file. 
> Done

#### (b) For each question, insert a new chunk and label the question number.
> Done

# Question 1

#### (a) Read the inflation consumer file into R and call it Inflation.df. 
```{r}
Inflation.df <- read.csv("inflation_consumer.csv")
```

#### (b) How many observations (rows) are in this dataset? How many variables (columns)? 
```{r}
num_rows <- nrow(Inflation.df)
num_cols <- ncol(Inflation.df)
num_rows
num_cols
```

#### (c) What is the mean value for the Year column? Does this value have any meaning? 
```{r}
mean_year <- mean(Inflation.df$Year) 
mean_year
```

#### (d) Determine which country has the lowest inflation and the corresponding year.
```{r}
min_inflation <- Inflation.df[which.min(Inflation.df$Inflation),]
min_inflation
```

#### (e) Determine which country has the highest inflation and the corresponding year.
```{r}
max_inflation <- Inflation.df[which.max(Inflation.df$Inflation),]
max_inflation
```


# Question 2

#### (a) Create a new data frame that contains all countries with inflation over 7.00 and save it to Sub Inflation.
```{r}
Sub_Inflation <- Inflation.df[Inflation.df$Inflation > 7.00,]
```

#### (b) Determine the average of the inflation column and save this to a variable called avg.inflation.
```{r}
avg.inflation <- mean(Sub_Inflation$Inflation)
avg.inflation
```

#### (c) Use Sub Inflation to determine the number of country counts with inflation over 15.00 and print out the value.
```{r}
num_overfifteen_countries <- nrow(Sub_Inflation[Sub_Inflation$Inflation > 15.00,])
num_overfifteen_countries 
```

#### (d) Create a vector named Inflation Status which contains two levels: High and Low. All countries in Sub Inflation with inflation over 15.00 should be labeled as High, otherwise named as Low.
```{r}
Sub_Inflation$Inflation_Status <- ifelse(Sub_Inflation$Inflation > 15, "High", "Low")
```

#### (e) Use the Inflation Status vector to print out the number of country counts with Inflation over that 15.00. This number should be the same as in part (c).
```{r}
num_high_inflation_status_countries <- sum(Sub_Inflation$Inflation_Status == "High")
num_high_inflation_status_countries
```


# Question 3

#### Once you make sure all the code works in the R markdown file, knit it to either an HTML or Word file. Make sure the file contains all answers to the questions. Then open the knitted file and print it as a PDF file. The name of the file should be Stat123 Lab03 YourLastName.pdf. Then submit the pdf file to the appropriate Brightspace folder.
> Done